### 2020-03-06 chapters/background.tex
Two-Stacks, DABA
Aggregation on data types

Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each window update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing.

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element insertion or eviction from the window, the aggregation value is updated, in contrast to RFS which would recompute the entire aggregation. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations. To illustrate, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the total sum plus the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed this far minus the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable.

@@

\begin{table}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/aggregation-operations-properties.png}
      \caption[Aggregation properties]{Aggregation properties. Checkmarks (\checkmark), crosses ($\times$), and question marks (?) indicate a property is true for all, false for all, or false for some of a given group of like operations, respectively}
      \label{tbl:aggregations-properties}
    \end{center}
\end{table}

@@

>>>>

Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each window update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing. Figure \ref{fig:pseudo-rfs} shows pseudocode for RFS \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.45]{figures/rfs.png}
      \caption{Recalculate-From-Scratch algorithm}
      \label{fig:pseudo-rfs}
    \end{center}
\end{figure}

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element insertion or eviction from the window, the aggregation value is updated, in contrast to RFS which would recompute the entire aggregation. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations. To illustrate, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the total sum plus the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed this far minus the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable. Figure \ref{fig:pseudo-soe} shows pseudocode for SOE \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.45]{figures/soe.png}
      \caption{Subtract-On-Evict algorithm}
      \label{fig:pseudo-soe}
    \end{center}
\end{figure}

@@

%TODO: change the sketch-like from the aggregator to the aggregation: bloom filter to membership query, CMS to frequency estimation and HLL to cardinality estimation
\begin{table}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/aggregation-operations-properties.png}
      \caption[Aggregation properties]{Aggregation properties. Checkmarks (\checkmark), crosses ($\times$), and question marks (?) indicate a property is true for all, false for all, or false for some of a given group of like operations, respectively}
      \label{tbl:aggregations-properties}
    \end{center}
\end{table}

@@

\section{Aggregations on data types}

Some aggregations are only possible on certain data types. For instance, computing an average on a set of strings is not possible. In this Section we present the data types we will work with and the set of possible aggregations to compute on them.

We consider four data types: \textit{numericals}, \textit{strings}, \textit{dates} and \textit{categoricals}. Numerical variables can be measured. When compared to classical programming data types, for example those of the C programming language, numericals comprise variable types \textit{short}, \textit{long}, \textit{int}, \textit{float}, \textit{double} and valid combinations of the previous --- \textit{e.g.} \textit{long int}. On the other hand, Strings are sequences of characters. The \textit{string} type is the most generic data type in that any other type can be converted to a string. Every variable, for example \textit{x}=3, can be turned into a string \textit{s}, for example \textit{s}="3". Similarly, a date is a sequence of characters and thus a string. However, despite being a string, it has semantics unlike a regular one. Knowing the date format, one can extract the day, month and year. For that reason we make the distinction between the \textit{string} and \textit{date} data types. Lastly, a \textit{categorical} is a variable that takes one of a relatively small and known set of possible values. A variable is assigned to a particular category based on its properties. The categorical variable is also a sequence of characters and therefore a string variable. The main difference between the categorical data type and the string is that we know the set of possible labels the categorical variable might take and this affects the type of computable aggregations.

On a set of numerical variables, we can compute sums, minima, maxima, averages, standard deviations and counts (of zeros, infinities and others). When operating on a set of strings, we can aggregate the collection into a concatenation of strings, a count, a membership query, cardinality.

### 2020-03-06 chapters/conclusion.tex
Aggregation on data types

We introduced background concepts on large scale data processing, windowing techniques applied to data streams and sliding window aggregations. We studied algorithms for change detection in data streams, pattern mining, anomaly detection and sliding window aggregation algorithms, specifically a subset of these named probabilistic data structures.

%%hypothesis validated?
The monitoring of data streams in real-time is applicable across many domains and applications. To the best of our knowledge, we claim that there has not been a real attempt at doing so in a lightweight way.

%%future work
Preliminary work shows that there exist constant space and time complexity algorithms to aggregate data, thus reducing memory usage. We intend to research more about lightweight sliding window aggregations and sliding implementations of probabilistic data structures. We also intend to search for new methods of summarizing data that do not necessarily explicitly work in a sliding window fashion. After identifying a set of algorithms that meet our requirements, the focus will be on implementing the best theoretical solution and benchmark our streaming system using data sets from the financial fraud space versus previously performed batch analysis.

%%summary of contributions
The main contribution of this Thesis will be the developed system code as well as a set of tests that measure the system's usefulness based on several parameters, such as recall, precision, memory used and latency.

>>>>

Yeeeeeeetttttt

### 2020-03-06 chapters/sota.tex
Aggregation on data types

@@

\iffalse
%TODO hash improvement probably better declared outside of bloom
An improvement to this structure was defined in \cite{Kirsch-Better-Bloom} where the authors attempt to reduce the overhead of bloom filters operations such as adding a new element or answering a query by focusing on the hash functions. The proposed method is based on hashing literature. It has been shown that multiple hash functions can be created from the combination of only two hash functions. This means that a new hash function \textit{g} can be produced based on  two other independent and uniform existing hash functions, \textit{h1} and \textit{h2}. The new hash function will map elements to a universe ranging from \textit{0} to \textit{p-1} and will be defined as \textit{$g(x) = h1(x) + ih2(x) mod p$}. Since the only nontrivial computations performed by a Bloom filter are the evaluations of pseudo-random hash functions, any reduction in the required number of pseudo-random hash functions yields a reduction in the time required to add an element or query a Bloom filter.

Bloom filters are not the only approximate aggregators that use hash functions as a basis of their algorithm. All of the studied probabilistic data structures do so. Therefore, the technique proposed is beneficial in HyperLogLogs, Count-Min Sketches and Bloom filters.
\fi

@@

\section{Sliding Window Aggregations with Probabilistic Data Structures}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space. These structures aggregate an endless stream of data. However, when working under a sliding window context there is the need not only to insert new events but to evict old ones. Probabilistic data structures aggregate all incoming data but have no way of expiring old elements from the aggregation state.

The probabilistic data structures studied can be implemented under generic sliding window aggregation algorithms like Recalculate-From-Scratch or DABA. However, doing so means we lose the main advantage of using these approximate aggregators: the constant space complexity. For example, cardinality estimation of a sliding window can be done employing the HyperLogLog heuristic of estimating cardinality as $2^\textit{p}$, with \textit{p} as the maximum position seen for the leftmost 1-bit, using DABA \ref{sec:daba} for the computation of the value \textit{p} per sliding window. With this approach, despite computing the aggregation state in constant time, we have a linear in window size space complexity solution. Thus, in this Section, we explore sliding window implementations of the previously discussed probabilistic data structures that keep the space and time complexity constant.

@@

\subsection{Sliding Window Cardinality Estimation and the Sliding HyperLogLog}

The HyperLogLog (HLL) aggregator presented in Section \ref{sec:hyper-log-log} will only be useful to us if it can be implemented efficiently in a sliding window fashion. To that end, we analyze the work of Chabchoub et al. that adapts the original HLL aggregator from Flajolet et al. to work under a sliding window framework  \cite{Chabchoub-Sliding-HLL}.

The HLL aggregator estimates on the set's cardinality assuming it is close to $2^\textit{p}$, where \textit{p} is the maximum position seen for the leftmost 1-bit in the binary values of all of the set items' hashes. In practice, all the aggregator has to compute is \textit{p}, the maximum value between a set of positions. Thus, the main idea behind Sliding HyperLogLog is to keep only the positions for the leftmost 1-bit that are possible future maximum values for \textit{p}.

The adaptation of the original HLL to work under a sliding window scenario requires the storage of additional information. The original HLL maintained \textit{m} integer counters, each storing the maximum for the leftmost 1-bit position of the analyzed binary strings.  The counter held the maximum value for all the seen binary strings so far and there was no way to expire old items. To allow eviction, the proposed data structure is very similar to the original \textit{m} integer counters but keeps a list of values per bucket and associates a timestamp to each.

The stochastic averaging process of HLL remains unchanged: incoming items are hashed and distributed across \textit{m} buckets. However, the authors propose keeping a short list of \textit{packets} per bucket instead of a single integer count. A \textit{packet} is defined as a pair <\textit{t\textsubscript{i}, p\textsubscript{i}}> where \textit{t\textsubscript{i}} is the arrival timestamp of the packet and \textit{p\textsubscript{i}} the position of the leftmost 1-bit in the binary hash representation of the associated value. Packets in the list are possible maxima over future windows. Because of that, the authors name this list a \textit{"List of Future Possible Maxima (LFPM)"}.

\subsection*{Element Insertion \& Eviction}
Similarly to the original HLL, incoming elements are hashed and the first \textit{n} bits of the string are used to map the new element to one of the \textit{m} buckets. The first \textit{n} bits are then discarded and the bit-pattern observable \textit{p} is computed from the remaining string.

@@

\subsection*{Sliding Window Cardinality Query}

Estimation of cardinality is done similarly to the original HLL cardinality estimation algorithm. To estimate the cardinality at a timestamp \textit{t}, select packets that are in the window --- \textit{i.e.} packets <t\textsubscript{i}, p\textsubscript{i}> from the LFPM where t\textsubscript{i} > \textit{t} - WindowSize. This filtering is necessary since the proposed algorithm evicts old packets only at element insertion. However, it would be possible to do so when estimating the cardinality or even periodically. From the filtered list of packets, compute the highest p\textsubscript{i} among them. Do this for each bucket and in the end compute the harmonic mean between all values, just like the original HLL. The result will be the averaged bit-pattern observable value \textit{p} and the cardinality is estimated to be $2^\textit{p}$.

\subsection*{Time complexity analysis}
When performing insertions or evictions, the algorithm requires scanning through the Lists of Future Possible Maxima (LFPMs) to delete old packets or packets that will never be window maxima. Similarly, we need to scan the lists when estimating the cardinality to filter for packets within our window. Denote \textit{L\textsubscript{n}} as the mean size in packets for each LFPM. There are \textit{m} buckets, each with a list. To scan through all lists we need to analyze \textit{m L\textsubscript{n}} elements.

The authors prove that \textit{L\textsubscript{n}} has an upper bound of \textit{ln(n/m)}, with \textit{n} as the exact cardinality of the set and \textit{m} as the number of buckets used. Hence, in the worst-case scenario, we need to analyze \textit{m ln(n/m)} elements. This means that implementing a Sliding HLL using Lists of Future Possible Maxima has \textit{O(ln(n))} time complexity, given the number of buckets \textit{m} is constant.

\subsection*{Space complexity analysis}
The memory used in the Sliding HLL aggregator implemented with Lists of Future Possible Maxima (LFPMs) amounts to the total size of the \textit{m} lists. Assuming 4 byte timestamps and 1 byte bit-pattern observables for each packet, the aggregator uses \textit{(4+1) L\textsubscript{n} m} bytes.

Given \textit{ln(n/m)} as the upper bound for \textit{L\textsubscript{n}}, this sliding implementation of the HLL aggregator consumes up to \textit{(4+1) ln(n/m) m} bytes of memory. Considering that the number of buckets \textit{m} is constant, the space complexity of a Sliding HLL implementation with LFPM is \textit{O(ln(n))}. Thus, the growth of the window in size is followed by a logarithmic increase in used memory.

@@

\subsection*{Applicability to our Hypothesis}
Implementing a Sliding HLL using Lists of Future Possible Maxima (LFPMs) shows logarithmic time and space complexity while maintaining the same accuracy as the original HLL approximate aggregator. Hence, this is a valid approach to test our Hypothesis.

>>>>

\textbf{TODO: Add table with SWAG algorithms, time and space complexity and applicable restrictions on the aggregation}

@@

@@

\iffalse
%TODO: Worth to show? HLL uses only one hash, only CMS and Bloom benefit from this.
\subsection{Hash improvement}
An improvement to this structure was defined in \cite{Kirsch-Better-Bloom} where the authors attempt to reduce the overhead of bloom filters operations such as adding a new element or answering a query by focusing on the hash functions. The proposed method is based on hashing literature. It has been shown that multiple hash functions can be created from the combination of only two hash functions. This means that a new hash function \textit{g} can be produced based on  two other independent and uniform existing hash functions, \textit{h1} and \textit{h2}. The new hash function will map elements to a universe ranging from \textit{0} to \textit{p-1} and will be defined as \textit{$g(x) = h1(x) + ih2(x) mod p$}. Since the only nontrivial computations performed by a Bloom filter are the evaluations of pseudo-random hash functions, any reduction in the required number of pseudo-random hash functions yields a reduction in the time required to add an element or query a Bloom filter.

Bloom filters are not the only approximate aggregators that use hash functions as a basis of their algorithm. All of the studied probabilistic data structures do so. Therefore, the technique proposed is beneficial in Count-Min Sketches and Bloom filters.
\fi

\section{Sliding Window Aggregations with Probabilistic Data Structures}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space. These structures aggregate an endless stream of data. However, when working with sliding windows there is the need not only to insert new events but to evict old ones. Probabilistic data structures aggregate all incoming data but have no way of expiring old elements from the aggregation state.

The probabilistic data structures studied can be implemented under generic sliding window aggregation algorithms like Recalculate-From-Scratch or DABA. However, doing so means we lose the main advantage of using these approximate aggregators: the constant space complexity. For example, cardinality estimation of a sliding window can be done employing the HyperLogLog heuristic of estimating cardinality as $2^\textit{p}$, with \textit{p} as the maximum position seen for the leftmost 1-bit, using DABA \ref{sec:daba} for the computation of \textit{p} per sliding window. With this approach, despite computing the aggregation state in constant time, we have a linear in window size space complexity solution. Thus, in this Section, we explore sliding window implementations of the previously discussed probabilistic data structures that keep the space and time complexity at the very least logarithmic.

@@

\subsection{Sliding HyperLogLog}

The HyperLogLog (HLL) aggregator presented in Section \ref{sec:hyper-log-log} will only be useful to us if it can be implemented efficiently in a sliding window fashion. The HLL aggregator estimates the set's cardinality assuming it is close to $2^\textit{p}$, where \textit{p} is the maximum position seen for the leftmost 1-bit in the binary values of all of the set items' hashes. In practice, all the aggregator has to compute is \textit{p}, the maximum value of all observables. Stochastic averaging is applied by emulating the effects of \textit{m} experiments using \textit{m} buckets to store \textit{m} bit-pattern observables.

\subsubsection{Sliding HyperLogLog using Lists of Future Possible Maxima} \label{sec:hll-lfpm}

The work of Chabchoub et al. adapts the original HLL aggregator from Flajolet et al. to work under a sliding window framework  \cite{Chabchoub-Sliding-HLL}.
The main idea behind this implementation is to replace every HLL counter with a \textit{"List of Future Possible Maxima (LFPM)"}.

\subsubsection*{Data Structure}

The adaptation of the original HLL to work under a sliding window scenario requires the storage of additional information. The original HLL maintained \textit{m} integer counters, each storing the maximum for the leftmost 1-bit position of the analyzed binary strings.  The counter held the maximum value for all the seen binary strings so far but there was no way to expire old items. To allow eviction, the proposed data structure is very similar to the original \textit{m} integer counters but keeps a list of values per bucket and associates a timestamp to each.

The stochastic averaging process of HLL remains unchanged: incoming items are hashed and distributed across \textit{m} buckets. However, the authors propose keeping a short list of \textit{packets} per bucket instead of a single integer count. A \textit{packet} is defined as a pair <\textit{t\textsubscript{i}, p\textsubscript{i}}> where \textit{t\textsubscript{i}} is the arrival timestamp of the packet and \textit{p\textsubscript{i}} the position of the leftmost 1-bit in the binary hash representation of the associated value. Packets are only kept in the list if they are possible maxima over future windows. Because of that, the authors name this list a \textit{"List of Future Possible Maxima (LFPM)"}.

\subsubsection*{Element Insertion \& Eviction}
Similarly to the original HLL, incoming elements are hashed and the first \textit{n} bits of the string are used to map the new element to one of the \textit{m} buckets. The first \textit{n} bits are then discarded and the bit-pattern observable \textit{p} is computed from the remaining string.

@@

\subsubsection*{Sliding Window Cardinality Query}

Estimation of cardinality is done similarly to the original HLL cardinality estimation algorithm. To estimate the cardinality at a timestamp \textit{t}, from all \textit{m} buckets and all \textit{m} lists, select packets that are in the window --- \textit{i.e.} packets <t\textsubscript{i}, p\textsubscript{i}> where t\textsubscript{i} > \textit{t} - WindowSize. This filtering is necessary since the proposed algorithm evicts old packets only at element insertion and at the time of the query some packets may no longer belong to the window. Next, from the filtered set of packets, for each bucket and its list, compute the maximum p\textsubscript{i} value. In the end, compute the harmonic mean between all values, just like the original HLL. The result will be the averaged bit-pattern observable value \textit{p} and the cardinality is estimated to be $2^\textit{p}$.

\subsubsection*{Time complexity analysis}
When performing insertions or evictions, the algorithm requires scanning through the Lists of Future Possible Maxima (LFPMs) to delete old packets or packets that will never be window maxima. Similarly, we need to scan the lists when estimating the cardinality to filter for packets within our window. Denote \textit{L\textsubscript{n}} as the mean size in packets for each LFPM. There are \textit{m} buckets, each with a list. To scan through all lists we need to analyze \textit{m L\textsubscript{n}} elements.

The authors prove that \textit{L\textsubscript{n}} has an upper bound of \textit{ln(n/m)}, with \textit{n} as the cardinality of the set and \textit{m} as the number of buckets used. Hence, in the worst-case scenario, we need to analyze \textit{m ln(n/m)} elements. This means that a Sliding HLL implementation using Lists of Future Possible Maxima has \textit{O(ln(n))} time complexity, given the number of buckets \textit{m} is constant.

\subsubsection*{Space complexity analysis}
The memory used in the Sliding HLL aggregator implemented with Lists of Future Possible Maxima (LFPMs) amounts to the total size of the \textit{m} lists. Assuming 4 byte timestamps and 1 byte bit-pattern observables for each packet, the aggregator uses \textit{(4+1) L\textsubscript{n} m} bytes.

The lists are dynamic thus they shrink and grow at run time. However, \textit{ln(n/m)} was proved to be an upper bound for \textit{L\textsubscript{n}}. Hence, this sliding implementation of the HLL aggregator consumes up to \textit{(4+1) ln(n/m) m} bytes of memory. Considering that the number of buckets \textit{m} is constant, the space complexity of a Sliding HLL implementation with LFPM is \textit{O(ln(n))}.

@@

\subsubsection*{Applicability to our Hypothesis}
Implementing a Sliding HLL using Lists of Future Possible Maxima (LFPMs) shows logarithmic time and space complexity while providing estimates with the same accuracy as the original HLL approximate aggregator. However, the memory consumption is dynamic due to the variable size of the lists.

Despite the fact memory consumption dynamically changes during runtime, this is a valid approach to test our Hypothesis due to the worst-case scenario being a logarithmic space complexity. However, in Section \ref{sec:hll-drv} we analyze an implementation that is not dynamic in memory during run time making it a preferred approach.

\subsubsection{Sliding HyperLogLog using Distance Recorder Vectors} \label{sec:hll-drv}

In Section \ref{sec:hll-lfpm} we have analyzed a Sliding HLL implementation where each of the \textit{m} buckets holds a list of values (LFPM) instead of a single integer count. However, the LFPM structure is dynamic and its size varies at run time.

The approach proposed by Xu uses a smaller amount of memory that does not change at run time \cite{Xu-hll-sliding-drv}. The author proposes a special counter: a \textit{"Distance Recorder Vector (DRV)"}.

\subsubsection*{Data Structure}
Much like the original HLL and the LFPM Sliding implementation, the Distance Recorder Vector (DRV) approach uses \textit{m} buckets. However, instead of integer counts or lists, each of the buckets stores a DRV.

A DRV is composed by \textit{$\lceil ln(n/m) \rceil$} records, with \textit{n} as the cardinality of the set. Notice that \textit{n/m} is the average number of distinct elements each bucket \textit{m} is expected to process and that a binary string of size \textit{$\lceil ln(n/m) \rceil$} can represent \textit{n/m} distinct elements. Also recall that the bit-pattern observable \textit{p} used by the HLL is the position of the leftmost 1-bit. In a binary string of size \textit{$\lceil ln(n/m) \rceil$}, \textit{p} will range from 1 to \textit{$\lceil ln(n/m) \rceil$}. The DRV is indexed exactly this way, from 1 to \textit{$\lceil ln(n/m) \rceil$}, and each bit-pattern observable \textit{p} will be used as an index in this vector of records. Essentially, each \textit{p} will have an associated record in a DRV of a bucket.

Initially, all records in the DRV will be set to be infinity or a large number.

\subsubsection*{Element Insertion}
When an element is inserted, the same procedure as in the original HLL is followed: the element is hashed and \textit{p} is computed from the hashed binary representation. Denote the bit-pattern observable of the new element as \textit{p\textsubscript{new}}. For the DRV of that bucket, increment the value of all records by one. Next, set position \textit{p\textsubscript{new}} in the DRV to 0 (DRV[\textit{p\textsubscript{new}}] = 0).

\subsubsection*{Sliding Window Cardinality Query} 
Element eviction is not performed. Instead, the query algorithm takes into account that some records in some of the DVR's might not be valid anymore. To estimate the sliding window cardinality is to find the value \textit{p}.

To find the value \textit{p} for a bucket, scan the DRV of that bucket backward. Start at the last element --- positioned at \textit{$\lceil ln(n/m) \rceil$} --- and scan towards the head of the vector. For each record, check if it is active in the current window. To do so, check if its value is less than the size of the window. If it is, the record is active. Otherwise, the record is inactive.

If the record is inactive continue scanning backward. If the record is active, then the position or index currently being scanned is the value of \textit{p} for that bucket. If all records are inactive, the bit-pattern observable \textit{p} is 0.

Stochastic averaging is done just like in the original HLL: compute the harmonic mean of all \textit{p\textsubscript{m}} values from all \textit{m} buckets. The cardinality of the sliding window is then estimated to be $2^\textit{p}$.

\subsubsection*{Time complexity analysis}
When inserting an element, determining its bucket, hashing the element and computing \textit{p} is done in constant time. However, all the  \textit{$\lceil ln(n/m) \rceil$} records must have their value increased by one and this means iterating over the DRV. Setting the element of the DRV indexed by \textit{p} to 0 is done in constant time as well.

Estimating cardinality is done by scanning the DRV backward until an active record is found. In the worst-case scenario, this means scanning all \textit{$\lceil ln(n/m) \rceil$} records.

Both inserting an element and estimating sliding window cardinality require a scan through \textit{$\lceil ln(n/m) \rceil$} records. This gives us a time complexity of \textit{O(\textit{ln(n)})}, given the number of buckets \textit{m} is constant.

\subsubsection*{Space complexity analysis}
This approach still uses \textit{m} buckets but each one contains a Distance Recorder Vector (DRV). A DRV is a vector of \textit{$\lceil ln(n/m) \rceil$} elements. Hence, \textit{$m \lceil ln(n/m) \rceil$} elements must be kept in memory, giving us a space complexity of \textit{O(ln(n))} with \textit{n} as the cardinality of the set.

\subsubsection*{Applicability to our Hypothesis}
Despite having the same time and space complexities, the Sliding HLL DRV approach uses a smaller amount of memory that does not change during runtime, while the LFPM requires heap resizing because the lists dynamically shrink and grow. Hence, the DRV approach is a better fit to integrate our lightweight real-time solution.
